\part{Extras}

% Things that were in previous parts but were put here because the parts
% were already too long


\section{Introduction}

\subsection{Design Rules}

\begin{frame}{Design Rules (1/4)}{General rules}
  \begin{block}{General rules}
    \begin{enumerate}
    \item
      \CCLang's evolution must be driven by real problems.
    \item
      Don't get involved in a sterile quest for perfection.
    \item
      \CCLang must be useful \emph{now}.
    \item
      Every feature must have a reasonably obvious implementation.
    \item
      Always provide a transition path.
    \item
      \CCLang is a language, not a complete system.
    \item
      Provide comprehensive support for each supported style.
    \item
      Don't try to force people to use a specific programming style.
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}{Design Rules (2/4)}{Design support rules}
  \begin{block}{Design support rules}
    \begin{enumerate}
    \item
      Support sound design notions.
    \item
      Provide facilities for program organization.
    \item
      Say what you mean.
    \item
      All features must be affordable.
    \item
      It is more important to allow a useful feature than to prevent every misuse.
    \item
      Support composition of software from separately developed parts.
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}{Design Rules (3/4)}{Language-technical rules}
  \begin{block}{Language-technical rules}
    \begin{enumerate}
    \item
      No implicit violations of the static type system.
    \item
      Provide as good support for user-defined types as for built-in types.
    \item
      Locality is good.
    \item
      Avoid order dependencies.
    \item
      If in doubt, pick the variant of a feature that is easiest to teach.
    \item
      Syntax matters (often in perverse ways).
    \item
      Preprocessor usage should be eliminated.
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}{Design Rules (4/4)}{Low-level programming support rules}
  \begin{block}{Low-level programming support rules}
    \begin{enumerate}
    \item
      Use traditional (dumb) linkers.
    \item
      No gratuitous incompatibilities with C.
    \item
      Leave no room for a lower-level language below C++ (except assembler).
    \item
      What you don’t use, you don’t pay for (zero-overhead rule).
    \item
      If in doubt, provide means for manual control.
    \end{enumerate}
  \end{block}
\end{frame}
