\part{Procedural programming}
\label{part:proc}

\section{Procedural programming}

% definition and declaration

\subsection{Arguments}

\subsubsection{Default arguments}

\begin{frame}{Default arguments}{}
  \begin{definition}[Default arguments]
    \strong{Default arguments} are used in calls where \emph{trailing} arguments are missing.
  \end{definition}

  \begin{example}[Default arguments]
    \sourceinput{snippets/default_arguments.cc}
  \end{example}
\end{frame}

\subsubsection{Variadic arguments}

\begin{frame}{Variadic arguments}{}
  \begin{definition}[Variadic arguments]
    \strong{Variadic arguments} are used when a function accepts any number of arguments. They are indicated by a parameter of the form \lstinline!...! (ellipsis) that must appear last in the parameter list of the function declaration.
  \end{definition}

  \begin{block}{Remarks}
    \begin{itemize}
    \item
      The ellipsis can be noted \lstinline!...! (\CCLang style) or \lstinline!, ...! (C style)
      \smallskip
      \sourceinput{snippets/ellipsis.cc}
    \item
      To be able to access the arguments, the function must have at least one named parameter before the ellipsis
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Accessing variadic arguments}{}
  \begin{block}{Accessing variadic arguments}
    The \lstinline!<cstdarg>! header provides facilities for accessing the variadic arguments:
    \begin{itemize}
    \item
      The \lstinline!va_list! type holds the information for accessing the variadic arguments
    \item
      The \lstinline!va_start! macro enables access to the variable arguments
    \item
      The \lstinline!va_arg! macro accesses the next variadic function argument
    \item
      The \lstinline!va_end! macro ends traversal of the variadic function arguments
    \end{itemize}
  \end{block}

  \begin{block}{Default argument promotions}
    \begin{itemize}
    \item
      \lstinline!nullptr_t! is converted to \lstinline!void*!
    \item
      \lstinline!float! is converted to \lstinline!double!
    \item
      \lstinline!bool!, \lstinline!char!, \lstinline!short! are converted to \lstinline!int!
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Example of variadic arguments}{}
  \begin{example}
    \sourceinput{snippets/variadic_arguments.cc}
  \end{example}
\end{frame}


\subsection{Overloading}

\subsubsection{Function overloading}

\begin{frame}{Function overloading}{}
  \begin{definition}[Function overloading]
    \strong{Function overloading} is the ability to create multiple functions with the same name with different implementations. Two overloaded function must differ by the type of their parameters.
  \end{definition}

  \begin{block}{Overload resolution}
    When a function is overloaded, overload resolution is the process to select the function that will be called.
    \begin{itemize}
    \item
      Determination of the set of candidate functions after name lookup and template argument deduction
    \item
      Determination of the set of viable functions after examining arguments and parameters
    \item
      Choice of the best viable function
    \end{itemize}
    Overload resolution may fail and result in a compilation error.
  \end{block}
\end{frame}

\begin{frame}{Function overloading}{}
  \begin{example}[Function overloading]
    \sourceinput{snippets/overload.cc}
  \end{example}
\end{frame}

\subsubsection{Operator overloading}

\begin{frame}{Operator overloading}{}
  \begin{block}{Operator overloading}
    Operator overloading is a special case of function overloading. Operators that can be overloaded are:
    \begin{itemize}
    \item
      Arithmetic operators: \lstinline!+!, \lstinline!-!, \lstinline!*!, \lstinline!/!, \lstinline!\%!, \lstinline!\^!, \lstinline!\&!, \lstinline!|!, \lstinline!\~!, \lstinline!<<!, \lstinline!>>!
    \item
      Increment and decrement operators: \lstinline!++!, \lstinline!--!
    \item
      Logical operators: \lstinline+!+, \lstinline!&&!, \lstinline!||!
    \item
      Assignment operators: \lstinline!=!, \lstinline!+=!, \lstinline!-=!, \lstinline!*=!, \lstinline!/=!, \lstinline!\%=!, \lstinline!\^=!, \lstinline!\&=!, \lstinline!|=!, \lstinline!<<=!, \lstinline!>>=!
    \item
      Comparisons operators: \lstinline!==!, \lstinline+!=+, \lstinline!<!, \lstinline!>!, \lstinline!<=!, \lstinline!>=!, \lstinline!<=>!\Since{20}
    \item
      Access operators: \lstinline!->!, \lstinline!->*!, \lstinline![]!
    \item
      Special operators: \lstinline!,! (comma), \lstinline!()! (call)
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Overloaded operators}{}
  \begin{block}{Overloaded operators}
    \begin{center}
      \scriptsize
      \begin{tabular}{|c|c|c|l|}
        \hline
        Expression & Member function & Free function & Example \\
        \hline
        \lstinline!@a! & \lstinline!(a).operator@()! & \lstinline!operator@(a)! & \lstinline+!std::cin+ \\
                      &                             &                          & $\to$ \lstinline+std::cin.operator!()+ \\
        \hline
        \lstinline!a @ b! & \lstinline!(a).operator@(b)! & \lstinline!operator@(a,b)! & \lstinline!std::cout << 42! \\
                          &                              &                            & $\to$ \lstinline!std::cout.operator<<(42)! \\
        \hline
        \lstinline!a = b! & \lstinline!(a).operator=(b)! &                            & \lstinline!std::string s; s = "abc";! \\
                          &                              &                            & $\to$ \lstinline!std::string.operator=("abc")! \\
        \hline
        \lstinline!a(b)! & \lstinline!(a).operator()(b)! &                            & \lstinline!std::random_device r; auto n = r();! \\
                        &                               &                            & $\to$ \lstinline!r.operator()()! \\
        \hline
        \lstinline!a[b]! & \lstinline!(a).operator[](b)! &                            & \lstinline!std::map<int, int> m; m[1] = 2;! \\
                        &                               &                            & $\to$ \lstinline!m.operator[](1)! \\
        \hline
        \lstinline!a->!  & \lstinline!(a).operator->()!  &                            & \lstinline!auto p = std::make_unique<S>(); p->bar()! \\
                        &                               &                            & $\to$ \lstinline!p.operator->()! \\
        \hline
        \lstinline!a@!   & \lstinline!(a).operator@(0)!  & \lstinline!operator@(a,0)! & \lstinline!auto i = v.begin(); i++! \\
                        &                               &                            & $\to$ \lstinline!i.operator++(0)! \\
        \hline
      \end{tabular}
    \end{center}
  \end{block}
\end{frame}

\begin{frame}{Restrictions on overloaded operators}{}
  \begin{block}{Restrictions on overloaded operators}
    \begin{itemize}
    \item
      The following operators can \strong{not} be overloaded:
      \begin{itemize}
      \item
        \lstinline!::! (scope resolution)
      \item
        \lstinline!.! (member access)
      \item
        \lstinline!.*! (member access through pointer to member)
      \item
        \lstinline!?:! (ternary conditional)
      \end{itemize}
    \item
      New operators such as \lstinline!**!, \lstinline!<>!, or \lstinline!&|! can \strong{not} be created
    \item
      The overloads of operators \lstinline!&&! and \lstinline!||! lose short-circuit evaluation
    \item
      The overload of operator \lstinline!->! must either return a raw pointer, or return an object (by reference or by value) for which operator \lstinline!->! is in turn overloaded
    \item
      It is \strong{not} possible to change:
      \begin{itemize}
      \item
        the precedence of operators
      \item
        the grouping of operators
      \item
        the number of operands of operators
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Operator overloading}{}
  \begin{example}[Operator overloading]
    \sourceinput{snippets/overload_operator.cc}
  \end{example}
\end{frame}


\subsection{Parameter passing}

% https://isocpp.github.io/CppCoreGuidelines/CppCoreGuidelines#Rf-conventional

\begin{frame}{Properties of types}{}
  \begin{block}{Properties of types}
    Types may be of three kinds regarding their ease of use in functions:
    \begin{enumerate}
    \item
      Easy types:
      \begin{itemize}
      \item
        Cheap to copy: \lstinline!int! and all fundamental types, \lstinline!std::tuple<bool, int*>!, \ldots
      \item
        Impossible to copy: \lstinline!std::unique_ptr<T>!, \ldots
      \end{itemize}
    \item
      Medium types:
      \begin{itemize}
      \item
        Cheap to move: \lstinline!std::string!, \lstinline!std::vector<T>!, \ldots
      \item
        Quite cheap to move: \lstinline!BigData!, \lstinline!std::array<std::vector<T>,N>!, \ldots
      \item
        Unknown cost: \lstinline!unknown::Type!, \ldots
      \end{itemize}
    \item
      Hard types:
      \begin{itemize}
      \item
        Expensive to move: \lstinline!std::array<BigData,N>!, \ldots
      \end{itemize}
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}{Output only values}{}
  \begin{block}{Output only values}
    For output only values, prefer:
    \begin{itemize}
    \item
      Return values for easy and medium types \\
      \lstinline!X f();!
%       \begin{itemize}
%       \item
%         In case of multiple return values, use a \lstinline!struct! or \lstinline!std::tuple! \\
%         \lstinline!std::tuple<X,Y> f();!
%       \end{itemize}
    \item
      Output parameter (reference to non-const) for hard types \\
      \lstinline!void f(X& out);!
    \end{itemize}
  \end{block}

  \begin{example}[Output only values]
    \sourceinput{snippets/parameter_passing_out.cc}
  \end{example}
\end{frame}

\begin{frame}{Input only values}{}
  \begin{block}{Input only values}
    For input only values, prefer:
    \begin{itemize}
    \item
      Passing by value for easy types \\
      \lstinline!void f(T in);!
    \item
      Passing by const reference for medium and hard types
      \lstinline!void f(const T& in);!
    \end{itemize}
  \end{block}

  \begin{example}[Input only values]
    \sourceinput{snippets/parameter_passing_in.cc}
  \end{example}
\end{frame}

\begin{frame}{Input/Output values}{}
  \begin{block}{Input/Output values}
    For input/output values, prefer:
    \begin{itemize}
    \item
      Passing by reference to non-const for any type \\
      \lstinline!void f(T& inout);!
    \end{itemize}
  \end{block}

  \begin{example}[Input/Output values]
    \sourceinput{snippets/parameter_passing_inout.cc}
  \end{example}
\end{frame}


\begin{frame}{Special case: copy of a parameter}{}
  \begin{block}{Copy of a parameter}
    In the case a function wants a copy of a parameter of medium type, the function can be overloaded to handle two situations:
    \begin{itemize}
    \item
      \lstinline!void f(const T& in);!
      \begin{itemize}
      \item
        The function will be used for lvalues
      \item
        The function will make a deep copy of the argument
      \end{itemize}
    \item
      \lstinline!void f(T&& in);!
      \begin{itemize}
      \item
        The function will be used for rvalues
      \item
        The function will move from the argument
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}


\subsection{Argument-dependent lookup}

% http://en.cppreference.com/w/cpp/language/adl

\begin{frame}{Argument-dependent lookup}{}
  \begin{definition}[Argument-dependent lookup]
    \strong{Argument-dependent lookup} (ADL) is the set of rules for looking up the unqualified function names in function-call expressions, including implicit function calls to overloaded operators. These function names are looked up in the namespaces of their arguments in addition to the scopes and namespaces considered by the usual unqualified name lookup.
  \end{definition}
\end{frame}

\begin{frame}{Argument-dependent lookup}{}
  \begin{example}[Argument-dependent lookup]
    \sourceinput{snippets/adl.cc}
  \end{example}
\end{frame}


\subsection{\texttt{constexpr} functions}

\begin{frame}{\texttt{constexpr} functions}{}
  \begin{block}{\texttt{constexpr} functions}
    A \lstinline!constexpr! functions must satisfy the following requirements\Since{14}:
    \begin{itemize}
    \item
      Its return type must be a literal type
    \item
      Each of its parameter must be a literal type
    \item
      The function body must \strong{not} contain:
      \begin{itemize}
      \item
        an \lstinline!asm! declaration
      \item
        a \lstinline!goto! statement
      \item
        a statement with a label other than \lstinline!case! and \lstinline!default!
      \item
        a definition of a variable of non-literal type
      \item
        a definition of a variable of static or thread storage duration
      \item
        a definition of a variable for which no initialization is performed
      \end{itemize}
    \end{itemize}

    A \lstinline!constexpr! function can be used to compute values at compile time.
  \end{block}

  \begin{block}{Remark}
    In \CC{11}, a \lstinline!constexpr! function could only have exactly one return statement!
  \end{block}
\end{frame}

\begin{frame}{\texttt{constexpr} functions}{}
  \begin{example}[\texttt{constexpr} functions]
    \sourceinput{snippets/constexpr_function.cc}
  \end{example}
\end{frame}


\subsection{User-defined literals}

\begin{frame}{Literal suffix}{}
  \begin{block}{Literal suffix}
    A \strong{literal suffix} is appended to a literal to specify a type for the literal:
    \begin{itemize}
    \item
      For integers: \lstinline!u!/\lstinline!U! for \lstinline!unsigned!, \lstinline!l!/\lstinline!L! for \lstinline!long! or \lstinline!ll!/\lstinline!LL! for \lstinline!long long!
    \item
      For floats: \lstinline!f!/\lstinline!F! for \lstinline!float!, \lstinline!l!/\lstinline!L! for \lstinline!long double!
    \end{itemize}
  \end{block}

  \begin{example}[Literal suffix]
    \sourceinput{snippets/literal_suffix.cc}
  \end{example}
\end{frame}

\begin{frame}{User-defined literals}{}
  \begin{block}{User-defined literals}
    \strong{User-defined literals} are literals with a user-defined suffix. They allow integer, floating-point, character, and string literals to produce objects of user-defined type by defining a user-defined suffix.
    \begin{itemize}
    \item
      The suffix is an identifier
    \item
      The suffix introduced by a program must begin with an underscore
    \item
      The standard library suffixes do not begin with underscores
    \end{itemize}
  \end{block}

  \begin{example}[User-defined literals]
    \sourceinput{snippets/user-defined_literals.cc}
  \end{example}
\end{frame}

\begin{frame}{Literal operator}{}
  \begin{block}{Literal operator}
    The function called by a user-defined literal is known as \strong{literal operator}. It is a function like any other with a name of the form:

    {
      \hfill\lstinline[mathescape]!operator "" $\;id$!\hfill
    }

    Only the following parameter lists are allowed on literal operators:
    \begin{itemize}
    \item
      \lstinline!(const char *)!
    \item
      \lstinline!(unsigned long long int)!
    \item
      \lstinline!(long double)!
    \item
      \lstinline[mathescape]!($character$)!
      \begin{itemize}
      \item
        where $character$ is one of \lstinline!char!, \lstinline!char8_t!\Since{20}, \lstinline!char16_t!, \lstinline!char32_t!
      \end{itemize}
    \item
      \lstinline[mathescape]!(const $\;character\;$ *, std::size_t)!
      \begin{itemize}
      \item
        where $character$ is one of \lstinline!char!, \lstinline!char8_t!\Since{20}, \lstinline!char16_t!, \lstinline!char32_t!
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Literal operator}{}
  \begin{example}[Literal operator]
    \sourceinput{snippets/literal_operator.cc}
  \end{example}
\end{frame}

\begin{frame}{Literal operators in the standard library}{}
  \begin{block}{Literal operators in the standard library}
    \begin{itemize}
    \item
      \lstinline!operator""if!, \lstinline!operator""i!, \lstinline!operator""il! for \lstinline!std::complex!\Since{14}
    \item
      \lstinline!operator""h!, \lstinline!operator""min!, \lstinline!operator""s!, \lstinline!operator""ms!, \lstinline!operator""us!, \lstinline!operator""ns! for \lstinline!std::chrono::duration!\Since{14}
    \item
      \lstinline!operator""y! for \lstinline!std::chrono::year!\Since{20}
    \item
      \lstinline!operator""d! for \lstinline!std::chrono::day!\Since{20}
    \item
      \lstinline!operator""s! for \lstinline!std::string!\Since{14}
    \item
      \lstinline!operator""sv! for \lstinline!std::string_view!\Since{17}
    \end{itemize}

    They are all defined in inline namespaces inside \lstinline!std::literals!
  \end{block}
\end{frame}

\subsection{Example: \texttt{string\_id}}

\begin{frame}{Example: \texttt{string\_id}}{}
  \begin{block}{Specifications for \texttt{string\_id}}
    \begin{itemize}
    \item
      Define a \lstinline!string_id! that is an integer that represents a string
      \begin{itemize}
      \item
        Need for a hash function: FNV-1a
      \end{itemize}
    \item
      Have a literal operator for creating a \lstinline!string_id! from a string
    \item
      Possibly do all this at compile-time
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Implementation of \texttt{string\_id}}{}
  \begin{example}[\texttt{string\_id}]
    \small
    \sourceinput{snippets/string_id.cc}
  \end{example}
\end{frame}


\begin{frame}{Use of \texttt{string\_id}}{}
  \begin{example}[\texttt{string\_id}]
    \small
    \sourceinput{snippets/string_id_use.cc}
  \end{example}
\end{frame}






% mangling

% copy elision
